Methodology for Assessing the State of the Practice for Domain X

### 2021-10-22 StateOfPractice/Methodology/Methodology.tex
Final version of Methodology paper, next step to prepare the arXiv version

\hypersetup{
colorlinks=true,     % false: boxed links; true: colored links
linkcolor=red,       % color of internal links (change box color with
%linkbordercolor)
citecolor=blue,      % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan        % color of external links
}

@@

\title{Methodology for Assessing the State of the Practice for Domain X} 
\author{Spencer Smith}{McMaster University, Canada}{smiths@mcmaster.ca}{}{}
\author{Jacques Carette}{McMaster University, Canada}{carette@mcmaster.ca}{}{}
\author{Olu Owojaiye}{McMaster University, Canada}{owojaiyo@mcmaster.ca}{}{}
\author{Peter Michalski}{McMaster University, Canada}{michap@mcmaster.ca}{}{}
\author{Ao Dong}{McMaster University, Canada}{donga9@mcmaster.ca}{}{}

\authorrunning{Smith et al.}  \Copyright{Spencer Smith and Jacques Carette and
Olu Owojaiye and Peter Michalski and Ao Dong}

@@

\begin{abstract}
	To improve software development methods and tools for research software, we
	first need to understand the current state of the practice.  Therefore, we
	have developed a methodology for assessing the state of the software
	development practices for a given research software domain.  The methodology
	is applied to one domain at a time in recognition that software development in
	different domains is likely to have adopted different best practices.
	Moreover, providing a means to measure different domains facilitates
	comparison of development practices between domains.  For each domain we wish
	to answer questions such as: 
  \begin{inparaenum}[i)]
    \item What artifacts (documents, code, test cases, etc.) are present?
    \item What tools are used?
    \item What principles, process and methodologies are used?
    \item What are the pain points for developers?
    \item What actions are used to improve qualities like maintainability and
    reproducibility?
  \end{inparaenum} 
  To answer these questions, our methodology prescribes the following steps: 
  \begin{inparaenum}[i)] 
    \item Identify the domain;
    \item Identify a list of candidate software packages;
    \item Filter the list to a length of about 30 packages;
    \item Gather source code and documentation for each package;
    \item Collect repository related data on each software package, like number
    of stars, number of open issues, number of lines of code;
    \item Fill in the measurement template (the template consists of 108
    questions to assess 9 qualities (including the qualities of installability,
    usability and visibility));
    \item Interview developers (the interview consists of 20 questions and takes
    about an hour);
    \item Rank the software using the Analytic Hierarchy Process (AHP); and,
    \item Analyze the data to answer the questions posed above.
  \end{inparaenum}
  A domain expert should be engaged throughout the process, to ensure that
  implicit information about the domain is properly represented and to assist
  with conducing an analysis of the commonalities and variabilities between the
  30 selected packages.  Using our methodology, spreadsheet templates and AHP
  tool, we estimate (based on our experience with using the process) the time to
  complete an assessment for a given domain at \total{totHours} person hours.
\end{abstract}

@@

Within this document the following typographic conventions have been adopted:
\begin{inparaenum}[i)] 
  \item \textcolor{red}{Red} text denotes a link to sections of this document. 
  \item \textcolor{cyan}{Cyan} text denotes a URL link.
  \item \textcolor{blue}{Blue} text denotes a citation.
\end{inparaenum}

@@

\begin{enumerate}
\item Identify the domain (X). (Section~\ref{SecIdentifyDomain})
\item Ask the Domain Expert to create a top ten list of software packages in the
  domain. (Section~\ref{SecDomainExpert})
\item Meet with the Domain Expert to brief them on the overall objective,
  research proposal, research questions, measurement template, and developer
  interviews. (Section~\ref{SecDomainExpert}) \label{StepMeeting}
\item Identify the broad list of candidate software packages in the domain.
  (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages list.
  (Section~\ref{SecInitialFilter})
\item Review software list with Domain Expert. (Section~\ref{SecDomainExpert})
\item Domain Analysis (with the help of the Domain Expert).
(Section~\ref{SecDomainAnalysis})
\item Ask Domain Expert to vet domain analysis.
  (Section~\ref{SecDomainExpert})
\item Gather source code and documentation for each prospective software
  package.
\item Collect repository based information. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecMeasureTemplate})
\item Survey developers. (Section~\ref{SecSurveyQuestions})
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
% \item Identify a short list of top software packages, typically four to six, for
%   deeper exploration according to the AHP rankings of the measurements.
\item Ask Domain Expert to vet AHP ranking. (Section~\ref{SecDomainExpert})
%\item Rank short list. (Section~\ref{SecRankShortList})
\item Answer research questions (from Section~\ref{SecResearchQuestions}) and
document answers.
\end{enumerate}

@@

\begin{figure}[h!]
	\begin{center}
	  \includegraphics[width=1.0\textwidth]{measurement_template}
	  \caption{Excerpt of the Top Section of the Measurement Template (Summary Information)}
	  \label{measurement_template_image}
	\end{center}
  \end{figure}

@@

As in \citet{SmithEtAl2016}, Virtual machines (VMs) are used to provide an
optimal testing environments for each package. VMs were used because it is
easier to start with a fresh environment without having to worry about existing
libraries and conflicts. Moreover, when the tests are complete the VM can be
deleted, without any impact on the host operating system. The most significant
advantage of using VMs is to level the playing field. Every software install
starts from a clean slate, which removes ``works-on-my-computer`` errors. When
filling in the measurement template spreadsheet, the the details for each VM
should be noted, including hypervisor and operating system version.

@@

The Analytical Hierarchy Process (AHP) is a decision-making technique that can
be used when comparing multiple options by multiple criteria. In our work AHP is
used for comparing and ranking the software packages of a domain using the
quality scores that are gathered in the Measurement Template
(Appendix~\ref{SecGradingTemplate}). AHP performs a pairwise analysis between
each of the 9 quality options for each of the 30 software packages.  This
results in a matrix, which is used to generate an overall score for each
software package for the given criteria. \cite{SmithEtAl2016} shows how AHP is
applied to ranking software based on quality measures. We have developed a
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{tool}
for conducting this process. The tool includes an AHP JAR script and a
sensitivity analysis JAR script that is used to ensure that the software package
rankings are appropriate with respect to the uncertainty of the quality scores.
The
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{README
file} outlines the requirements for, and configuration and usage of, the JAR
scripts. The JAR scripts, source code, and required libraries are located in the
same folder as the README file.

@@

- usability and modifiability experiments

Usability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})

Modifiability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})

- not just for research software

- use it to inform research on Drasil - what artifacts, what process dev
currently use, how could the new process promoted by Drasil fit into the current
development processes?

- meta analysis of collected data.  Already have GIS, mesh generator.  Currently
working on medical image analysis and LBM.

@@

- summary

@@

View the analytic results by open [output path]/index.html with any browser or
other software supporting HTML web page format.
\begin{figure}[h!]
\includegraphics[scale=0.4]{git-stats-demo.JPG}
\end{figure}

@@

\begin{figure}[h!]
\includegraphics[scale=0.5]{scc-demo.JPG}
\end{figure}

@@

\def\arraystretch{1.4}
\begin{tabular}{p{14cm}}
	\hline	
	\textbf{Maintainability}\\
	\hline
	What is the current version number? (number)\\
	Is there any information on how code is reviewed, or how to contribute?
	({yes$^*$, no})\\
	Are artifacts available? (List every type of file that is not a code file –
	for examples please look at the ‘Artifact Name’ column of
	https://gitlab.cas.mcmaster.ca/SEforSC/se4sc/-/blob/git-svn/GradStudents/Olu/ResearchProposal/Artifacts\_MiningV3.xlsx)
	({yes$^*$, no, unclear}) $^*$list via string\\
	What issue tracking tool is employed? (set of {Trac, JIRA, Redmine, e-mail,
	discussion board, sourceforge, google code, git, BitBucket, none, unclear,
	other$^*$}) $^*$ given via string\\
	What is the percentage of identified issues that are closed? (percentage)\\
	What percentage of code is comments? (percentage)\\
	Which version control system is in use? ({svn, cvs, git, github, unclear,
	other$^*$}) $^*$ given via string\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline		
	\textbf{Reusability}\\
	\hline
	How many code files are there? (number)\\
	Is API documented? ({yes, no, n/a})\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline		
\end{tabular}

@@

% \def\arraystretch{1.22}
% \begin{tabular}{p{14cm}}
% \hline
% \textbf{Repo Metrics (Measured via git_stats)}\\
% \hline
% Number of text-based files. (number)\\
% Number of binary files. (number)\\
% Number of total lines in text-based files. (number)\\
% Number of total lines added to text-based files. (number or {not git repo})\\
% Number of total lines deleted from text-based files. (number or {not git repo})\\
% Number of total commits. (number or {not git repo})\\
% Numbers of commits by year in the last 5 years. (Count from as early as possible if the project is younger than 5 years.) (list of numbers or {not git repo}\\
% Numbers of commits by month in the last 12 months. (list of numbers or {not git repo})\\
% \hline
% \end{tabular}

% \def\arraystretch{1.22}
% \begin{tabular}{p{14cm}}
% \hline
% \textbf{Repo Metrics (Measured via scc)}\\
% \hline
% Number of text-based files.	(number)\\
% Number of total lines in text-based files.	(number)\\
% Number of code lines in text-based files.	(number)\\
% Number of comment lines in text-based files.	(number)\\
% Number of blank lines in text-based files.	(number)
% \hline
% \end{tabular}

% \def\arraystretch{1.22}
% \begin{tabular}{p{14cm}}
% \hline
% \textbf{Repo Metrics (Measured via GitHub)}\\
% \hline
% Number of stars.	(number or {n/a})
% Number of forks.	(number or {n/a})
% Number of people watching this repo.	(number or {n/a})
% Number of open pull requests.	(number or {n/a})
% Number of closed pull requests.	(number or {n/a})
% \hline
% \end{tabular}

@@

Introduction.  Roadmap.  Anything about traceability to research questions.

Information about these interview questions:  This gives you an idea what I
would like to learn about the development of {domain} software. Interviews will
be one-to-one and will be open-ended (not just “yes or no” answers). Because of
this, the exact wording may change a little. Sometimes I will use other short
questions to make sure I understand what you told me or if I need more
information when we are talking such as: “So, you are saying that ...?), to get
more information (“Please tell me more?”), or to learn what you think or feel
about something (“Why do you think that is...?”).

\subsection{Information about the developers and users} \citep{Jegatheesan2016}
\begin{enumerate}
\item Interviewees' current position/title? degrees?
\item Interviewees' contribution to/relationship with the software?
\item Length of time the interviewee has been involved with this software?
\item How large is the development group?
\item Do you have a defined process for accepting new contributions into your team?
\item What is the typical background of a developer?
\item What is your estimated number of users? How did you come up with that estimate?
\item What is the typical background of a user?
\end{enumerate}

@@

Square brackets are used for traceability to qualities and relevant research
questions found in Section ...

\begin{enumerate}
\item Currently, what are the most significant obstacles in your development
process?
\item How might you change your development process to remove or reduce these
obstacles?
\item How does documentation fit into your development process? Would improved
documentation help with the obstacles you typically face? [traceability -
research question 5b, visibility/transparency - research question 5i]
\item In the past, is there any major obstacle to your development process that
has been solved? How did you solve it?
\item What is your software development model? For example, waterfall, agile,
etc.
\item What is your project management process? Do you think improve this process
can tackle the current problem? Were any project management tools used?
\item Was it hard to ensure the correctness of the software? If there were any
 obstacles, what methods have been considered or practiced to improve the
 situation? If practiced, did it work? [correctness - research question 5e]
\item When designing the software, did you consider the ease of future changes?
 For example, will it be hard to change the structure of the system, modules or
 code blocks? What measures have been taken to ensure the ease of future changes
 and maintains? [maintainability - research question 5d, modifiability -
 research question 5c]
\item Provide instances where users have misunderstood the software. What, if
any, actions were taken to address understandability issues? [understandability
- research question 5f]
\item What, if any, actions were taken to address usability issues? [usability -
research question 5a]
\item Do you think the current documentation can clearly convey all necessary
knowledge to the users? If yes, how did you successfully achieve it? If no, what
improvements are needed? [unambiguity - research question 5g]
\item Do you have any concern that your computational results won't be
reproducible in the future? Have you taken any steps to ensure reproducibility?
[reproducibility - research question 5h]
\end{enumerate}

>>>>

\hypersetup{
colorlinks=true,     % false: boxed links; true: colored links
linkcolor=blue,      % color of internal links (change box color with linkbordercolor)
citecolor=blue,      % color of links to bibliography
filecolor=magenta,   % color of file links
urlcolor=cyan        % color of external links
}

@@

\title{Methodology for Assessing the State of the Practice for Domain X} 
\author{Spencer Smith}{McMaster University, Canada}{smiths@mcmaster.ca}{}{}
\author{Jacques Carette}{McMaster University, Canada}{carette@mcmaster.ca}{}{}
\author{Peter Michalski}{McMaster University, Canada}{michap@mcmaster.ca}{}{}
\author{Ao Dong}{McMaster University, Canada}{donga9@mcmaster.ca}{}{}
\author{Olu Owojaiye}{McMaster University, Canada}{owojaiyo@mcmaster.ca}{}{}

\authorrunning{Smith et al.}  \Copyright{Spencer Smith and Jacques Carette and
 Peter Michalski and Ao Dong and Olu Owojaiye}

@@

\begin{abstract}
	To improve software development methods and tools for research software, we
	first need to understand the current state of the practice.  Therefore, we
	have developed a methodology for assessing the state of the software
	development practices for a given research software domain.  The methodology
	is applied to one domain at a time in recognition that software development in
	different domains is likely to have adopted different best practices.
	Moreover, providing a means to measure different domains facilitates
	comparison of development practices between domains.  For each domain we wish
	to answer questions such as: 
  \begin{inparaenum}[i)]
    \item What artifacts (documents, code, test cases, etc.) are present?
    \item What tools are used?
    \item What principles, process and methodologies are used?
    \item What are the pain points for developers?
    \item What actions are used to improve qualities like maintainability and
    reproducibility?
  \end{inparaenum} 
  To answer these questions, our methodology prescribes the following steps: 
  \begin{inparaenum}[i)] 
    \item Identify the domain;
    \item Identify a list of candidate software packages;
    \item Filter the list to a length of about 30 packages;
    \item Gather source code and documentation for each package;
    \item Collect repository related data on each software package, like number
    of stars, number of open issues, number of lines of code;
    \item Fill in the measurement template (the template consists of 108
    questions to assess 9 qualities (including the qualities of installability,
    usability and visibility));
    \item Interview developers (the interview consists of 20 questions and takes
    about an hour);
    \item Rank the software using the Analytic Hierarchy Process (AHP); and,
    \item Analyze the data to answer the questions posed above.
  \end{inparaenum}
  A domain expert should be engaged throughout the process, to ensure that
  implicit information about the domain is properly represented and to assist
  with conducting an analysis of the commonalities and variabilities between the
  30 selected packages.  Using our methodology, spreadsheet templates and AHP
  tool, we estimate (based on our experience with using the process) the time to
  complete an assessment for a given domain at \total{totHours} person hours.
\end{abstract}

@@

Within this document the following typographic conventions have been adopted:
\begin{inparaenum}[i)] 
  \item \textcolor{red}{Blue} text denotes a link to sections of this document,
  including citations 
  \item \textcolor{cyan}{Cyan} text denotes an external URL link.
\end{inparaenum}

@@

\begin{enumerate}
\item Identify the domain (X). (Section~\ref{SecIdentifyDomain})
\item Ask the Domain Expert to create a top ten list of software packages in the
  domain. (Section~\ref{SecDomainExpert})
\item Meet with the Domain Expert to brief them on the overall objective,
  research proposal, research questions, measurement template, and developer
  interviews. (Section~\ref{SecDomainExpert}) \label{StepMeeting}
\item Identify the broad list of candidate software packages in the domain.
  (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages list.
  (Section~\ref{SecInitialFilter})
\item Review software list with Domain Expert. (Section~\ref{SecDomainExpert})
\item Domain Analysis (with the help of the Domain Expert).
(Section~\ref{SecDomainAnalysis})
\item Ask Domain Expert to vet domain analysis.
  (Section~\ref{SecDomainExpert})
\item Gather source code and documentation for each prospective software
  package.
\item Collect repository based information. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecMeasureTemplate})
\item Survey developers. (Section~\ref{SecSurvey})
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
% \item Identify a short list of top software packages, typically four to six, for
%   deeper exploration according to the AHP rankings of the measurements.
\item Ask Domain Expert to vet AHP ranking. (Section~\ref{SecDomainExpert})
%\item Rank short list. (Section~\ref{SecRankShortList})
\item Answer research questions (from Section~\ref{SecResearchQuestions}) and
document answers.
\end{enumerate}

@@

\begin{figure}[!ht]
	\begin{center}
	  \includegraphics[width=1.0\textwidth]{measurement_template}
	  \caption{Excerpt of the Top Section of the Measurement Template (Summary Information)}
	  \label{measurement_template_image}
	\end{center}
  \end{figure}

@@

As in \citet{SmithEtAl2016}, Virtual machines (VMs) are used to provide an
optimal testing environments for each package. VMs were used because it is
easier to start with a fresh environment without having to worry about existing
libraries and conflicts. Moreover, when the tests are complete the VM can be
deleted, without any impact on the host operating system. The most significant
advantage of using VMs is to level the playing field. Every software install
starts from a clean slate, which removes ``works-on-my-computer'' errors. When
filling in the measurement template spreadsheet, the the details for each VM
should be noted, including hypervisor and operating system version.

\section{Survey Developers} \label{SecSurvey}

In the previous state of the practice measurement process
\citep{SmithEtAl2018_arXivGIS, SmithEtAl2016, SmithEtAl2018}, we only based our
assessment on information available in on-line software repos.  However, this
approach meant we weren't able to learn about the development process, the
attitudes of the developers, the pain points for developers and how the
developers handle modifiability, reproducibility and usability.  Therefore, in
the reboot of the measurement process, we have explicitly added a stage for
interviewing research software developers.

We designed a list of 20 questions to guide our interviews, which can be found
in Appendix~\ref{SecSurveyQuestions}. Some questions are about the background of
the software, the development teams, the interviewees, and how they organize the
projects. We also ask about the developer's understandings of the users. Some
questions focus on the current and past difficulties, and the solutions the team
has found, or will try. We also discuss the importance and current situations of
documentation. A few questions are about specific software qualities, such as
maintainability, understandability, usability, and reproducibility. The
interviews are semi-structured based on the question list; we ask follow-up
questions when necessary. Based on our experience, the interviewees usually
bring up some exciting ideas that we did not expect, and it is worth expanding
on these topics.

Our methodology suggests requesting interviews with a developer from each of the
30 software package.  Requests for interviews are sent to all packages so as to
not cause a potential bias by singling out any subset of the list. Moreover,
since not every developer will agree to the interview request, asking 30 times
will typically yield a reasonable number of responses. In our experience, the
response rate is between 15\% and 30\%.  In some cases multiple developers from
the same project will agree to be interviewed. When sending out interview
requests, we recommend finding the contacts on the projects’ website, or code
repository, or publications, or the biographic pages of the teams’ institutions.
We send at most two interview request emails to a contact for each software
package.  Meeting will typically be held using on-line meeting software, like
Zoom or Teams.  This facilitates recording and automatic transcription of the
meetings.

The interviewees should follow a process where they can make informed consent.
The interviews should follow standard ethics guideline of asking for consent
before interviewing, recording, and including participant details in the report.
The interview process presented here was approved by the McMaster University
Research Ethics Board under the application number 
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/MACREM/Application.pdf}
{MREB\#: 5219}.

@@

The Analytical Hierarchy Process (AHP) is a decision-making technique that can
be used when comparing multiple options by multiple criteria. AHP focuses on
pair-wise comparisons between all options for all criteria.  The advantage of
pair-wise comparisons is that they facilitates a separation of concerns.  Rather
than worry about the entire problem, the decision maker can focus on one
comparison at a time.  In our work AHP is used for comparing and ranking the
software packages of a domain using the quality scores that are gathered in the
Measurement Template (Appendix~\ref{SecGradingTemplate}). AHP performs a
pairwise analysis between each of the 9 quality options for each of the 30
software packages.  This results in a matrix, which is used to generate an
overall score for each software package for the given criteria.
\cite{SmithEtAl2016} shows how AHP is applied to ranking software based on
quality measures. We have developed a
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{tool}
for conducting this process. The tool includes an AHP JAR script and a
sensitivity analysis JAR script that is used to ensure that the software package
rankings are appropriate with respect to the uncertainty of the quality scores.
The
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{README
file} outlines the requirements for, and configuration and usage of, the JAR
scripts. The JAR scripts, source code, and required libraries are located in the
same folder as the README file.

@@

As explained in the introduction (Section~\ref{SecIntroduction}), our eventual
goal is to improve the state of practice of software development, which requires
a baseline means for measuring the current state of the practice.  Now that we
have a methodology for assessing the state of practice for a given domain, the
next task is to complete the measurements for multiple domains.  Using the
previous and updated methodologies, we have measured the state of
practice for the following domains: Geographic Information Systems
\citep{SmithEtAl2018_arXivGIS}, Mesh Generators \citep{SmithEtAl2016},
Seismology software \citep{SmithEtAl2018}, Statistical software for psychology
\citep{SmithEtAl2018_StatSoft}, medical image analysis software \citep{Dong2021}
and for Lattice Boltzmann Method software \citep{Michalski2021}.  Future domains
for measurement could include finite element software, computational medicine,
machine learning, ordinary differential equation solvers, computer graphics,
stoichiometry, etc.

With the wealth of data from assessing the state of practice for multiple
domains, the next step is a meta-analysis.  We would look at how the different
domains compare. What lessons from one domain could be applied in other domains?
What (if any) differences exist in the pain points between domains?  Are there
differences in the tools, processes, and documentation between domains?

The current methodology is constrained by limited resources.  A 4 hour cap on
the measurement time for each software package limits what can be assessed.
Within this limit, we can't measure some important qualities, like usability and
modifiability.  In the future, we propose a more time-consuming process that
would capture these other quality measures.  To improve the feasibility, the
more time consuming measurements would not have to be completed for all 30
packages. Instead, a short list could be identified using the output of the AHP
ranking to select the top projects, or to select a sample of interesting
projects across the quality spectrum.

If we can add measures for modifiability and usability, we can start to measure
the quality impact of software development processes, tools and techniques.  For
instance, we have a project (called
\href{https://github.com/JacquesCarette/Drasil} {Drasil}
\citep{SzymczakEtAl2016}), which facilitates a development process that focuses
on knowledge capture, followed by generation of code and other software
artifacts from the captured knowledge.  To understand the advantages and
disadvantages of Drasil, we could measure its quality.  In particular, we would
like to see the impact of a generative process on the quality of usability and
modifiability.

\subsection{Measuring Usability in the Future}

In the future, we propose an experiment for assessing the usability of a given
software package.  Some initial thoughts on how this might be done are recorded
in this section.  To do the experiment we need an experimental subject, who will
be required to complete tasks with the software being studied.  The interaction
with the software will allow the study subject to experience the software's
usability.  The tasks for the subject to complete would vary by domain;
therefore, the tasks would be selected with the help of the Domain Expert.
Criteria for selecting candidate tasks are as follows:

\begin {enumerate}
\item Tasks should be executable for subjects with novice to intermediate
experience.
\item All tasks should take no more than one hour.
\item Tasks should include the basic/common use cases of the software package.
\item Include tasks that require sequential or hierarchical steps for
completion.
\end {enumerate}

Once the study subject has experience with the software, they will be in a
position to judge its usability.  We will measure the usability using a
standardized usability questionnaire, like the
\href{https://www.usabilitest.com/sus-pdf-generator} {System Usability Scale
questionnaire} or the
\href{https://uiuxtrend.com/pssuq-post-study-system-usability-questionnaire/}
{Post-Study System Usability Questionnaire}.

As a starting point for the experiment design, the procedure could be something
like the following:

\begin {enumerate}
\item Survey participants to collect pre-experiment data (background, experience
of subjects (especially with domain software)).
\item Participants perform tasks based on task defined by Domain Experts.
\item Observe the study subjects (take notes, record sessions (screen
recorder), watch for body languages and verbal cues).
\item Survey the study subjects to collect feedback (post-experiment interview),
complete usability questionnaire.
\item Prepare a summary report of the experimental results.
\end {enumerate}

\subsection{Measuring Modifiability in the Future}

The next experiment is designed to gather qualitative data regarding the
modifiability of each software package. This proposed experiment also involves
experimental subjects/participants, who in this case are asked to modify a set
of software packages.  The specific modifications requested will again depend on
the software domain.  In advance of the experiment the Domain Expert will be
asked the likely changes for software in this domain.  We emphasize likely
changes, instead of any changes because software cannot be designed so that
everything is equally easy to change \citep{parnas1986rational}.  The procedure
could be along the following lines:

\begin{enumerate}
	\item Domain Expert lists all likely changes that a developer
	might be asked to make in a software package in the domain.
	\item Survey study participants to collect pre-experiment data (background,
	experience of subjects (especially with domain software)).
	\item Participants perform modification tasks for likely changes on each
	software package being studied.
	\item Observe the study subjects (take notes, record sessions (screen
	recorder), watch for body languages and verbal cues).
	\item Record time needed to make the changes.	
	\item Confirm through testing that the modified software has the correct
	behaviour.
	\item Survey the study subjects to collect feedback (post-experiment interview).
	\item Prepare a summary report of the experimental results.
\end{enumerate}

The study subjects should make the same changes in multiple pieces of software.
The reporting of the results will focus more on the relative time differences
between the set of software packages, rather than the absolute time to make any
given change. To remove biases caused by the participants experience, the
different study subjects should use a different order as they go through the
list of software packages.  Details for this experiment still need to be
resolved, such as how to take the participants prior knowledge into account,
especially with respect to their programming experience.

@@

We have outlined a methodology for assessing the state of the practice for any
given research software domain.  (Although the scope of the current work has been
on research software, there is little in the methodology that is specific to
research software, except for the interview question related to the quality of
reproducibility.)  When applying the methodology to a given domain, we provide a
means to answer the following questions:
\begin{inparaenum}[i)]
\item What artifacts (documents, code, test cases, etc.) are present?
\item What tools are used?
\item What principles, process and methodologies are used?
\item What are the pain points for developers?
\item What actions are used to improve qualities like maintainability and
reproducibility?
\item What specific actions are taken to achieve the qualities of usability,
traceability, modifiability, maintainability, correctness, understandability,
unambiguity, reproducibility and visibility/transparency?
\item How does software designated as high quality by this methodology compare
  with top rated software by the community?
\end{inparaenum}

The methodology depends on the engagement of a Domain Expert.  The Domain
Expert's role is to ensure that the assessment is consistent with the culture of
the community of practitioners in the domain.  The Domain Expert also has an
important role to play with the domain analysis.  For each domain we 
conduct a domain analysis to look at the commonalities, variabilities and
parameters of variation, for the family of software in the domain.  The domain
analysis means that software can be compared not just based on its quality, but
also based on its functionality.

The methodology follows a systematic procedure that begins with identifying the
domain and ends with answering the research questions posed above.  In between
we collect an authoritative list of about 30 software packages.  For each
package in the list we fill in our measurement template.  The template consists
of repository related data (like number of open issues, number of lines of code,
etc.) and 108 measures/questions related to 9 qualities: installability,
correctness/verifiability, reliability, robustness, usability, maintainability,
reusability, understandability and visibility/transparency. Filling in the
template requires installing the software, running simple tests (like completing
the getting started instructions (if present)), and searching the code,
documentation and test files.

The data for each domain is used to rank the software package according to each
quality dimension using AHP.  The ranking is not intended to identify a single
best software package.  Instead the ranking is intended to provide insights on
the top set of software for each quality.  The top performers can be contrasted
with the lesser performers to gain insight into what practices in the domain are
working.  Deeper insight can be obtained by combining this data with the
interview data from asking each recruited developer 20 questions.

Combining the quantitative data from the measurement template with the interview
results, along with the domain experts knowledge, we can determine the current
state of the practice for domain X.  Using our methodology, spreadsheet
templates and AHP tool, we estimate (based on our experience with using the
process) the time to complete an assessment for a given domain at
\total{totHours} person hours.

@@

View the analytic results by open [output path]/index.html with any browser or
other software supporting HTML web page format.

\begin{figure}[!ht]
\includegraphics[scale=0.4]{git-stats-demo.JPG}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.5]{scc-demo.JPG}
\end{figure}

@@

\def\arraystretch{1.4}
\begin{tabular}{p{14cm}}
	\hline	
	\textbf{Maintainability}\\
	\hline
	What is the current version number? (number)\\
	Is there any information on how code is reviewed, or how to contribute?
	({yes$^*$, no})\\
	Are artifacts available? (List every type of file that is not a code file –
	for examples please look at the ‘Artifact Name’ column of
	https://gitlab.cas.mcmaster.ca/SEforSC/se4sc/-/blob/git-svn/GradStudents/Olu/ResearchProposal/Artifacts\_MiningV3.xlsx)
	({yes$^*$, no, unclear}) $^*$list via string\\
	What issue tracking tool is employed? (set of {Trac, JIRA, Redmine, e-mail,
	discussion board, sourceforge, google code, git, BitBucket, none, unclear,
	other$^*$}) $^*$ given via string\\
	What is the percentage of identified issues that are closed? (percentage)\\
	What percentage of code is comments? (percentage)\\
	Which version control system is in use? ({svn, cvs, git, Github, unclear,
	other$^*$}) $^*$ given via string\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline		
	\textbf{Reusability}\\
	\hline
	How many code files are there? (number)\\
	Is API documented? ({yes, no, n/a})\\
	Overall impression? ({1 .. 10})\\
	Additional comments? (can cover any metrics you feel are missing, or any other
	thoughts you have)\\
	\hline		
\end{tabular}

@@

@@

Interviews will be one-to-one and will be open-ended (not just “yes or no”
answers). Because of this, the exact wording of the questions may change a
little. If more information, or clarification, is needed during the
conversation, follow up questions will be asked, such as: ``So, you are saying
that ...?'', ``Please tell me more?'', or ``Why do you think that is?''

\subsection{Information about the developers and users}

These questions are based on questions used in \citet{Jegatheesan2016}

\begin{enumerate}
\item Interviewees' current position/title? degrees?
\item Interviewees' contribution to/relationship with the software?
\item Length of time the interviewee has been involved with this software?
\item How large is the development group?
\item Do you have a defined process for accepting new contributions into your team?
\item What is the typical background of a developer?
\item What is your estimated number of users? How did you come up with that estimate?
\item What is the typical background of a user?
\end{enumerate}

@@

Square brackets are used for traceability to the  relevant research
questions found in Section~\ref{SecResearchQuestions}.

\begin{enumerate}
\item Currently, what are the most significant obstacles in your development
process?
\item How might you change your development process to remove or reduce these
obstacles?
\item How does documentation fit into your development process? Would improved
documentation help with the obstacles you typically face? [research question 5b (traceability),
research question 5i (visibility/transparency)]
\item In the past, is there any major obstacle to your development process that
has been solved? How did you solve it?
\item What is your software development model? For example, waterfall, agile,
etc.
\item What is your project management process? Do you think improve this process
can tackle the current problem? Were any project management tools used?
\item Was it hard to ensure the correctness of the software? If there were any
 obstacles, what methods have been considered or practiced to improve the
 situation? If practiced, did it work? [research question 5e (correctness)]
\item When designing the software, did you consider the ease of future changes?
 For example, will it be hard to change the structure of the system, modules or
 code blocks? What measures have been taken to ensure the ease of future changes
 and maintains? [research question 5d (maintainability), 
 research question 5c (modifiability)]
\item Provide instances where users have misunderstood the software. What, if
any, actions were taken to address understandability issues? [research question
5f (understandability)]
\item What, if any, actions were taken to address usability issues? [
research question 5a (usability)]
\item Do you think the current documentation can clearly convey all necessary
knowledge to the users? If yes, how did you successfully achieve it? If no, what
improvements are needed? [research question 5g (unambiguity)]
\item Do you have any concern that your computational results won't be
reproducible in the future? Have you taken any steps to ensure reproducibility?
[research question 5h (reproducibility)]
\end{enumerate}
